---
title: "Elliptical Billiards: Triangular Orbits"
author: Dan S. Reznik
date: April, 2019
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: false
    theme: united
    df_print: paged
    highlight: tango
    code_folding: show
---

Source files available [here](https://github.com/dan-reznik/Elliptical-Billiards-Triangular-Orbits)

```{r,echo=F}
knitr::opts_chunk$set(
  cache=T,
  collapse=T,
  comment="#>",
  dpi=96,
  out.width='100%'
)
```

### Load packages

```{r,message=F}
library(tidyverse)
#library(gganimate)
# library(magrittr)
library(assertthat)
library(glue) # for string interpolation
library(magick) # for building animation
library(furrr) # for parallel processing
library(tictoc)# for timing
source("./util.R") # geom utilities, see repo
```

## Shape of error function

Note: goes to zero as $t$ approaches $\pi/2$, because ray-trace path is shrinking. 

```{r}
draw_error <- function(a,t) {
  p0 <- ell_p(a,t)
  n0 <- ell_n(a,t)
  df <- tibble(th=seq(-.99*pi/2,0,pi/500),
                   tr=map(th,tri_angle2,a=a,p=p0,norm=n0),
                   d2=map_dbl(tr,"d2")%>%sqrt,
                   d_adj=map_dbl(tr,"d_adj")) %>%
    select(-tr)
  
  df %>%
    gather("key","value",d2,d_adj) %>%
    ggplot(aes(th,value,group=key,color=key)) +
    geom_line()
}

draw_error(2,pi/4)
```

### Optimization functions

```{r}
# tri_angle_d2 <- function(th,a,p,norm) tri_angle2(th,a,p,norm)$d2
tri_angle_d_adj <- function(th,a,p,norm) tri_angle2(th,a,p,norm)$d_adj
opt_th <- function(a,p,norm,tol=1e-4)
optimize(tri_angle_d_adj,
         interval=c(-.99,-.05)*pi/2, # there's a direct back-and-forth hit at t in 8~9 degrees
         tol=tol,a=a,p=p,norm=norm)
# global
#opt_th_global <- function(a,p,norm,tol=1e-4) optim(c(-.99*pi/2,0),function(x) #tri_angle_d_adj(x,a=a,p=p,norm=norm),method="BFGS")
```

Calculate optimum $\theta$ for this triangle

```{r,include=F}
opt_th(2,ell_p(2,pi/4),ell_n(2,pi/4))
```

Investigate optimum $\theta$ for various values of $t$ from -90 to 0 degrees

```{r}
df_opt_th <- tibble(t=seq(-90,90,.5),
                        opt_th=map(t%>%to_rad,
                                   ~opt_th(2,ell_p(2,.x),ell_n(2,.x)))) %>%
  mutate(min=map_dbl(opt_th,"minimum"),
         obj=map_dbl(opt_th,"objective")) %>%
  select(-opt_th)

df_opt_th %>%
  gather("key","value",min,obj) %>%
  ggplot(aes(t,value,group=key,color=key)) +
  geom_line()
```

Some values are causing minimizer to be unstable

```{r}
summary(df_opt_th)
```

Values are between 8 and 9 degrees

```{r}
df_opt_th%>%filter(obj>.001)
```

```{r}
draw_error(2,8%>%to_rad)
```

### Get optimum triangle

```{r}
get_opt_tri <- function(t_deg,a) {
  t <- t_deg %>% to_rad
  p <- ell_p(a,t)
  n <- ell_n(a,t)
  opt_th <- opt_th(a,p,n)
  tri_angle_df(opt_th$minimum,a,p)
}
```

## Orbit perimeter is constant but area is not

```{r}
df_per_area <- tibble(t=seq(0,359,5),
                      tris=t%>%map(get_opt_tri,a=2),
                      sides=tris%>%map(get_tri_sides),
                      per=sides%>%map_dbl(get_tri_per),
                      area=sides%>%map_dbl(get_tri_area))

df_per_area %>%
  mutate(per) %>%
  gather(key,value,per,area) %>%
  ggplot(aes(t,value,group=key,color=key)) +
  geom_line(size=1) + 
  scale_y_log10(breaks=scales::log_breaks(10))
```

```{r,echo=F}
ggsave("pics/perimeter_area.png",width=7,height=3.5)
```

Fixed point on boundary, vary output angle wrt normal

```{r}
show_variable_angle <- function(a=2,t_deg=45,
                                th_min=-52,th_max=-50,th_step=.5) {
  t <- t_deg %>% to_rad
  p = ell_p(a,t)
  n = ell_n(a,t)
  n_tip = p + n*.1
  arrow15 <- arrow(angle = 15, type = "closed",length=unit(10,"points"))

  df_ell <- tibble(th = seq(th_min,th_max,th_step),
                       tri_angle = th %>% to_rad %>%
                         map(~tri_angle_df(.x,a,p,n))) %>%
    unnest(tri_angle) # expands to the side!
  
  # Animate
  
  df_ell %>%
    ggplot() +
    # background
    geom_point(x=0,y=0) +
    geom_path(aes(px,py),data=ell_poly(a)) +
    #geom_segment(x=p[1],y=p[2],
    #             xend=n_tip[1],yend=n_tip[2],
    #             color="#0000ff",arrow=arrow15) +
    # moving
    geom_path(aes(x=p_x,y=p_y,group=th)) +
    geom_point(aes(x=p_x,y=p_y,group=th),color="blue") +
    # labs(title = 'th: {th_range}') +
    geom_point(x=p[1],y=p[2],color="red") +
    coord_fixed() +
    facet_wrap(~th) # transition_manual(row)
}
show_variable_angle()
```

## Calculate incenter locus

Note: 150 degrees base vertex motion sweeps 360

```{r}
get_incenter_locus <- function(a,deg_step=1)
  seq(0,360,deg_step) %>%
  map(~attr(get_opt_tri(.x,a),"incenter")) %>%
  map_dfr(~tibble(x=.x[1],y=.x[2]))
```

Compute / retrieve incenter locus as data frame

```{r}
fname_incenter <- "data/incenter_locus.csv"
if (file.exists(fname_incenter)) {
  df_incenter <- read_csv(fname_incenter)
} else {
  df_incenter <- get_incenter_locus(2) %>%
    mutate_all(~round(.,4))
  # avoid multiple windings
  max_row <- df_incenter %>% 
    {which(.$x<0&.$y>0&lead(.$y)<0)} # y zero crossing
  df_incenter %>%
    head(max_row) %>%
    write_csv(fname_incenter)
}
```

## Show orbit triangle

```{r}
show_opt_tri <- function(t_deg,a,incenter_locus=NULL) {
  df_ell_bound <- ell_poly(a)
  df_ell_foci <- ell_foci(a)
  df_opt_tri <- get_opt_tri(t_deg,a)
  incenter <- attr(df_opt_tri,"incenter")

  arrow15 <- arrow(angle = 15, type = "closed",length=unit(10,"points"))  
  ggplot(df_opt_tri) +
    # fixed
    geom_point(x=0,y=0,shape=4) +
    geom_path(aes(px,py),df_ell_bound) +
    #geom_point(x=p0[1],y=p0[2]) +
    geom_point(aes(fx,fy),df_ell_foci) +
    geom_polygon(aes(p_x,p_y),
                 linetype="dotted",
                 color="blue",fill=NA) +
    geom_segment(aes(x=p_x,y=p_y,xend=ntip_x,yend=ntip_y),arrow=arrow15) +
    geom_point(aes(p_x,p_y),color="blue") +
    geom_point(aes(p_x,p_y),df_opt_tri%>%slice(1),color="red") +
    {
      if(is.null(incenter_locus))
        geom_blank()
      else
        geom_path(aes(x,y)
                  ,data=incenter_locus
                  ,color="green"
                  ,linetype=2
                  #,shape="."
        )  
    } + 
    geom_point(x=incenter[1],y=incenter[2],
               color="green",size=3) +
    coord_fixed() +
    labs(title="triangular orbit",
         subtitle=glue("a={a}, t={t_deg}°"),x="",y="") #+
    # theme(plot.margin=grid::unit(c(0,0,0,0), "mm"))
}
show_opt_tri(45,2,df_incenter)
```

## Compute animation

Save frames to files

```{r,eval=F,message=F}
fnames_png <- list.files(path = "./pics",
                         pattern = "tri_.*\\.png$",
                         full.names = T)
if (length(fnames_png)>0)
  file.remove(fnames_png)

tic()
plan("multiprocess")
deg_vec <- seq(0,360,1)
walk_vec <- deg_vec %>%
  future_iwalk(~{
    fname<-sprintf("./pics/tri_%04d.png",.y)
    p_tri<-show_opt_tri(.x,2,df_incenter)
    fname%>%ggsave(p_tri,dpi = 100)
    # fname%>%image_read%>%image_trim%>%image_write(fname)
  })
toc()
```

Combine frames into animated .gif

```{r,eval=F,message=F}
fnames_png <- list.files(path = "./pics",
                         pattern = "tri_.*\\.png$",
                         full.names = T)
fnames_png %>% 
  map(image_read) %>% # reads each path file
  map(image_trim) %>% # trims blank borders
  image_join() %>% # joins image
  # image_animate(fps=5) %>% # animates, can opt for number of loops
  # image_write_video("tris_a=2.mp4",framerate=10)
  image_write_gif("pics/tris_a=2.gif",delay=.1)

file.remove(fnames_png)
```

<!-- gganimate, buggy! -->

```{r,include=F,eval=F}
anim_opt_tri <- function(a) {
  df_ell_bound <- ell_poly(a)
  df_ell_foci <- ell_foci(a)
  df_opt_tri <- tibble(t=seq(0,359,1),
                           tri=t%>%map(get_opt_tri,a=a)) %>%
    unnest(tri)
  # View(df_opt_tri)
  # df_opt_tri
  df_opt_tri %<>%
    group_by(t) %>%
    mutate(pfirst_x=first(p_x),
           pfirst_y=first(p_y)) %>%
    ungroup

  arrow15 <- arrow(angle = 15, type = "closed",length=unit(10,"points"))
  df_opt_tri %>%
    ggplot(aes(x=p_x,y=p_y)) +
    # fixed
    geom_point(x=0,y=0,shape=4) +
    geom_path(aes(px,py),df_ell_bound) +
    #geom_point(x=p0[1],y=p0[2]) +
    geom_point(aes(fx,fy),df_ell_foci)  +
    geom_polygon(linetype="dotted",color="#0000ff",fill=NA) +
    geom_segment(aes(xend=ntip_x,yend=ntip_y),arrow=arrow15) +
    geom_point(color="#0000ff") +
    geom_point(aes(pfirst_x,pfirst_y),color="#ff0000") +
    coord_fixed() +
    labs(title = 't={frame_time%>%as.integer}, a={a}') +
    #labs(title="triangular orbit",
     #    subtitle=glue("a={a}, t: {frame_time}°"),x="",y="")
    transition_time(t) #+
    #ease_aes("linear")
}
anim_opt_tri(2)
```